Project 1
========================================================

### Loading and preprocessing the data

```{r echo=TRUE}
data<-read.csv("C:/Users/ppp/Documents/activity.csv",header=T)
head(data)
```


### Q1: What is mean total number of steps taken per day?

**Making a histogram of the total number of steps taken each day.**  
**Calculating and reporting the mean and median total number of steps taken per day.**

```{r fig.width=7, fig.height=6,echo=TRUE}
list1<-split(data$steps,data$date)
totalsteps<-as.numeric(sapply(list1,sum))
hist(totalsteps,main='The total number of steps taken each day',xlab='steps per day')
summary(totalsteps)
```

**From the above output, we find that the mean total number of steps taken per day is 10770.**
**The median total number of steps taken per day is 10760.**  
  
  
### Q2: What is the average daily activity pattern?

**Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).**

```{r fig.width=7, fig.height=6, echo=TRUE}
meansteps<-as.numeric(tapply(data$steps,data$interval,mean,na.rm=T))
interval<-as.numeric(unique(data$interval))
mean.data<-data.frame(mean.step=meansteps,interval=interval)
plot(mean.data[,2],mean.data[,1],type="l",main='Time series plot',xlab="intervals", ylab="mean steps")
mean.data[mean.data[,1]==max(mean.data[,1]),]
```

**5-minute interval "835", on average across all the days in the dataset, contains the maximum number of steps.**  


### Q3: Inputing Missing Values

**Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.**   
**How many missing values are there in the dataset?**

```{r echo=TRUE}
missing=numeric(3)
for (i in 1:3){
  missing[i]<-sum(is.na(data[,i])==TRUE)
}
sum(missing)
```

**There are 2304 missing values in the dataset.**    


**Using the mean for that 5-minute interval to fill in all of the missing values in the dataset.**  
**Creating a new dataset that is equal to the original dataset but with the missing data filled in.**

```{r echo=TRUE}
miss<-data[is.na(data$steps)==TRUE,]
nonmiss<-data[is.na(data$steps)==FALSE,]

miss.new<-miss

for (i in 1:2304){
  for (j in 1: 288){
    if ( miss.new[i,3] == as.integer(mean.data[j,2]) ){
      miss.new[i,1] <- mean.data[j,1] 
      break
    }
  }
}

data.new <- rbind(miss.new,nonmiss)
data.final<-data.new[order(data.new$date,data.new$interval),]
head(data.final)
```

**Making a histogram of the total number of steps taken each day.**  
**Calculating and reporting the mean and median total number of steps taken per day.**

```{r fig.width=7, fig.height=6,echo=TRUE}
list2<-split(data.final$steps,data.final$date)
totalsteps2<-as.numeric(sapply(list2,sum))
hist(totalsteps2,main='The total number of steps taken each day without missing values',xlab='steps per day')
summary(totalsteps2)
```

**The mean total number of steps taken per day is 10770.**  
**The median total number of steps taken per day is 10770.**  
**The mean is the same as the previous mean with missing values.**  
**The median is 10770,larger than 10760, the previous median with missing values.**  


### Q4: Are there differences in activity patterns between weekdays and weekends? 

**Creating a new factor variable in the dataset with two levels - "weekday" and "weekend".** 

```{r, echo=TRUE}
date <- as.Date(data.final[,2])
weekday<-weekdays(date)

weekde <- character(17568)
for (i in 1:17568){
  if (weekday[i] == "Saturday" || weekday[i]=="Sunday"){
    weekde[i] <- c("weekend")
  }
  else{
    weekde[i] <- c("weekday")
  }
}
data.final$weekde <- factor(weekde)

head(data.final)
```


**Making a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).**

```{r echo=TRUE}
data.split<-split(data.final, data.final$weekde)

data.weekday<-as.data.frame(data.split[1])
data.weekend<-as.data.frame(data.split[2])

table.weekday<-tapply(data.weekday$weekday.steps,data.weekday$weekday.interval,mean)
weekday.mean<-as.numeric(table.weekday)
weekday.interval<-as.numeric(rownames(table.weekday))
weekday.weekde<-rep("weekday",288)
weekday.final<-data.frame(mean.steps=weekday.mean,interval=weekday.interval,weekde=weekday.weekde)

table.weekend<-tapply(data.weekend$weekend.steps,data.weekend$weekend.interval,mean)
weekend.mean<-as.numeric(table.weekend)
weekend.interval<-as.numeric(rownames(table.weekend))
weekend.weekde<-rep("weekend",288)
weekend.final<-data.frame(mean.steps=weekend.mean,interval=weekend.interval,weekde=weekend.weekde)

week.final<-rbind(weekday.final,weekend.final)

# Lattice plotting methods
library(lattice)
xyplot( mean.steps ~ interval | weekde, data=week.final, type="l", layout=c(1,2) )

# Base plotting system
par(mfrow=c(2,1))
plot(weekend.final$interval,weekend.final$mean.steps,type="l",xlab='intervals',ylab='number of steps',main="Weekend")
plot(weekday.final$interval,weekday.final$mean.steps,type="l",xlab='intervals',ylab='number of steps',main="Weekday")
```
